The supernova-regulated ISM - III. Magnetic fields in the multi-phase gas

### 2015-03-21 paper.tex
CCE text edits

\begin{abstract}
Abstract to be left until the end. Words, words and more words to follow but first we need a paper. There will be methods, keywords and results hinted at here. 
\end{abstract}

\begin{keywords}
circumstellar matter -- infrared: stars.
\end{keywords}

@@

%FAG rephrase
The decomposition of the magnetic field into mean and fluctuating (random) 
field follows \MHD.
%end FAG
%When considering the mean-field decomposition of the magnetic field, the notation used by \citep{GSSFM13} is adopted.
Volume averaging with a Gaussian kernel, $G_{l}(\bvec{x}-\bvec{x}')$ is used to decompose the magnetic field, $\bvec{B}$, into mean (large-scale), $\bvec{B}_{l}$ and random (small-scale), $\bvec{b}_{l}$, fields. The decomposition is given by 
\begin{equation}
\bvec{B} = \bvec{B}_{l}+\bvec{b}_{l},\qquad \bvec{B}_{l} = \avg{B}. \label{decomp} 
\end{equation}
Gaussian smoothing of the magnetic field with kernel size $l$ is represented by
\begin{align}
\avg{B}&=\int_{V}\bvec{B}(\bvec{x}')G_{l}(\bvec{x}-\bvec{x}')d^{3}\bvec{x}',\\
G_{l}(\bvec{x})&=(2\pi l^2)^{-3/2}\exp[-\bvec{x}/(2l^2)],\nonumber
\label{gauss}
\end{align} 
where $l\approx50$pc is half the integral scale of random motions, as discussed
in \HD.  
reliminary analysis does not show significant sensitivity of the mean or random field to variations in $l$ within the range $0<l<100$pc. However, future work will consider analysis of isolated structures in the ISM. A more detailed analysis of length and correlation scales will be carried out. However, the current definition, $l=50$pc, is sufficient for the purposes of this paper. 
%\begin{table*}
 %\centering
 %\begin{minipage}{140mm}
  %\caption{Data on the RV Tauri stars detected by {\it IRAS}.}
  %\begin{tabular}{@{}llrrrrlrlr@{}}
  %\hline
  % Name     &            & \multicolumn{4}{c}{Flux density (Jy)%
 %\footnote{Observed by {\em IRAS}.}}\\
 % Variable & {\it IRAS} & 12$\,\umu$m & 25$\,\umu$m & 60$\,\umu$m
 %    & 100$\,\umu$m & Sp. & Period & Light- & $T_0\,(\rmn{K})$ \\
 %       &  &  &  &  &  & group & (d) & curve \\
 %       &  &  &  &  &  &       &     & type  \\
 %\hline
 %TW Cam & 04166$+$5719 & 8.27 & 5.62 & 1.82 & $<$1.73 & A & 85.6 & a & 555 \\
 %RV Tau & 04440$+$2605 & 22.53 & 18.08 & 6.40 & 2.52 & A & 78.9 & b & 460 \\
%\hline
%\end{tabular}
%\end{minipage}
%\end{table*}

@@

  As discussed in Section~\ref{subsect:params} the stability of the disc
  reduces the effectiveness of the SNe to generate and circulate hot gas, in
  the absence of SN clustering. 
  Hence Model~$\Omph$ has a slightly thicker disc, with less hot gas than 
  Model~\Op.
  On top of this the effect of the magnetic pressure in Model~$\Ompa$ is to 
  expand the thick disc even further and this is illustrated in 
  Fig.~\ref{fig:zrhom}a, where the horizontal averages of gas number density
  $n(z)$ are plotted against $z$ for both models. 
  The strong peak in the density at the mid-plane is evident for Model~$\Omph$
  (black, solid), while for Model~$\Ompa$ (blue, dashed) there is a broad
  plateau in the density, extending to $|z|\simeq300\pc$, where the mean 
  magnetic field is strongest.

@@

  The horizontal averages of the pressure are plotted in Fig.~\ref{fig:zrhom}b
  for both models. 
  There is a strong peak at the mid-plane in the turbulent pressure for the HD
  model (black, dashed), but a much weaker profile for the MHD model (blue,
  dash-dotted). 
  There are two peaks in the magnetic pressure (blue, dash-3dotted) near 
  $|z|\simeq200\pc$, which supports the extended density profile.
  Another possible effect, which might constrain the circulation of the hot
  gas, and hence enhance the pressure at the mid-plane, is the strong 
  horizontal orientation of the field.
  As mentioned in Section~\ref{sect:dyn} the periodic boundary conditions
  exclude a non-zero vertical component to the mean field, so the magnetic
  tension predominantly acts against the vertical flows.
  Some understanding of the multi-phase structure of the magnetised ISM is
  still possible from these models, but the extreme temperatures and densities
  are significantly under represented.
  To improve this in future work it will be desirable to allow unrestricted 
  evolution of vertical field and to apply realistic clustering of the 
  SNe to generate more supperbubbles (composite multiple SN remnants forming a
  single superstructure) or chimneys (plumes venting hot gas from the disc 
  towards the halo).

@@

Even though the visualisation is informative, it is necessary to determine whether the field lines are sensitive to the phases of the ISM. Consequently, we compare the characteristics of the field lines against straight lines launched in the positive $y$-direction, from the seed points used to compute the field lines. This test is devised using ideas from the field of stereology \citep{BJ04}, lineal analysis in particular. The straight lines used in the test case will have the same general characteristics as the field lines, since they both traverse the computational domain in the positive $y$-direction. This ensures that the lines are representative of the field lines being sampled. Despite having the same general characteristics, the straight lines should not be sensitive to the phases of the ISM; they will not prefer to stay in a particular phase, whilst avoiding others. Therefore, comparison of the PDFs of the observables sampled along the field lines and the straight lines, should identify sensitivity (or lack thereof) of the field lines to the phases of the ISM.\\

@@

The mean magnetic field is considered first, to explore the large-scale characteristics of the magnetic field, having extracted the small-scale fluctuations with Gaussian volume averaging.  Intuition suggests that the mean magnetic field will prefer to stay in the warm phase, which provides a less hostile environment for the magnetic field than the transonic, compressible turbulence of the hot phase. As discussed in Section \label{sect:3BB}, the magnetic field is very strong and well-aligned with the azimuthal ($y$) direction in the $|z|<0.6$kpc region, where the warm phase dominates. In the $0.6<|z|<1.0$kpc region, the hot phase dominates. Consequently, the magnetic field is not well aligned and much weaker than the magnetic field in the warm phase. The the turbulent flows in the $0.6<|z|<1.0$kpc region will have the highest kinetic energies throughout the ISM (CALCULATE THESE AVERAGE VALUES). Conversely, the magnetic field is very weak in the hot phase (Fig.~\ref{fig:b2dent}). Subsequently, the kinetic energy of the flows should be at least of a comparable order to the magnetic field energy in this region.  We expect this to cause distortions in the magnetic field (CCE: REFERENCE TAYLOR HERE). The lack of alignment and weakness of the magnetic field in this region (Fig. \ref{fig:b3box}) imply that the magnetic field is sensitive to the hot phase. The characteristics of the magnetic field lines further clarify the effect of the hot phase on the magnetic field.

@@

In Figure \ref{fig:fld_lines}, data from the snapshot at $t=1.4$ Gyr are used.
The white field lines have been seeded along the $xz$-axis at $y=-0.51$ kpc. 
The translucent gas inside the box is a representation of the log density field.
Density decreases from brighter to darker colours. 
Since the warm and hot phase represents approximately $\%99$ of the volume of the ISM gas, it is only useful for visualising the characteristics of the field line within the warm and hot phases.

In Figure \vref{fig:fld_lines}, the white field lines have been seeded along 
the $xz$-axis at $y\approx-0.51$ kpc. 
The translucent gas inside the box is a representation of the log density, 
$\log(n)$, field. 
Density decreases from brighter to darker colours. 
Since the warm and hot phase represents approximately $\%99$ of the fractional
volume of the ISM gas, it is only useful for visualising the characteristics 
of the field line within the warm and hot phases. 
We will present a preliminary result regarding the field lines within the 
cold phase in this paper. 
However, a more detailed discussion will be reported elsewhere. 
For descriptive purposes, the yellow shades of gas correspond to warm gas and 
the red shades correspond to the hot phase. 
We note that the field lines are generally aligned with the positive 
$y$-direction, especially in the warm phase and the field lines are smooth 
inside the warm gas. 
However, we notice significant differences in the characteristics of the 
field lines in the hot phase.

@@

Similarly, in the range $0.7<z<1.1$ kpc, there is a larger structure of hot 
gas, which spans the $xy$-plane. 
This structure appears to be part of rising hot gas, which is leaving the 
computational domain. As it leaves, it elongates the magnetic field lines 
until they are no longer connected inside the domain. 
Consequently, the the field lines are approximately perpendicular to the 
$y$-axis, where the magnetic field lines are strecthed and distorted by the 
hot phase.

\noindent 
Figure \ref{fig:pdfs_mean} shows the probability density functions (PDFs) of 
entropy sampled along field lines (solid lines) and straight lines (dashed
lines) for the snapshot at $t=1.4$ Gyr. 
The PDFs show that the magnetic field lines and straight lines have similar 
overall behaviour, due to their alignment in the $y$-direction. 
The alignment of the straight lines in the azimuthal ($y$) direction was 
chosen to ensure that the straight lines are representative of the general 
alignment of the magnetic field, without introducing any sensitivity to the 
multi-phase structure.

@@

Overall, the magnetic field lines traverse the ISM in the azimuthal direction. 
This trend is observed clearly in the warm phase, where the magnetic field is 
not distorted by the flows of the ISM gas. 
Since the field lines represent lines of magnetic flux density, the 
distortion of the magnetic field lines by the hot phase explains the weakness 
of the magnetic field observed in the hot phase. 
The stretching of the magnetic field lines effectively wrap the field lines 
around the hot gas, which increases the field line density on the boundary 
of the hot and warm gas. 
Consequently, field lines do not pass through the hot gas 
(Fig.~\ref{fig:fld_lines}) and we observe that the magnetic field in the hot 
phase is very weak and lacking alignment. 
This strongly suggests that the magnetic field resides in the warm phase.    
%-----------------------------------------------------------------------------
\begin{table}
 \centering
 %\begin{minipage}{180mm}
  \caption{Respective probabilities of finding the field and straight in the cold, warm and hot phases.}
  \label{table:mean_probs}
 % \begin{tabular*}{@{}c c c @{}}{\linewidth}{@{\extracolsep{\fill}}p{0.3\linewidth}p{0.3\linewidth}p{0.3\linewidth}@{}}
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}p{0.3\linewidth}p{0.3\linewidth}p{0.3\linewidth}@{}}
  \hline
   Probability& Mean field lines& Straight lines \\
 \hline
\\
$\mathcal P(\text{cold phase})$&$\ll1\%$ &$\ll1\%$  \\
   $\mathcal P(\text{warm phase})$& $0.79$&$0.89$   \\
   $\mathcal P(\text{hot phase})$&$0.20$ & $0.11$    \\
\hline
\end{tabular*}
%\end{minipage}
\end{table}
%-----------------------------------------------------------------------------

@@

%-----------------------------------------------------------------------------
\subsection{Curve fitting for PDFs of entropy}
%-----------------------------------------------------------------------------
\subsection{Correlation lengths of the magnetic fields}
\begin{table}[H]
 \centering
 %\begin{minipage}{180mm}
  \caption{Correlation lengths for the total, mean and random magnetic field in kpc.}
  \label{table:mean_probs}
 % \begin{tabular*}{@{}c c c @{}}{\linewidth}{@{\extracolsep{\fill}}p{0.3\linewidth}p{0.3\linewidth}p{0.3\linewidth}@{}}
\begin{tabular*}{\linewidth}{p{0.20\linewidth}p{0.20\linewidth}p{0.2\linewidth}p{0.2\linewidth}@{}}
  \hline
Correlation length &Total magnetic field& Mean magnetic field& Random magnetic field \\
 \hline
\\
$C$&$0.711$&$0.700$&$0.241$ \\
$C_x$& $0.061$&$0.045$&$0.065$   \\
$C_y$&$0.380$ & $0.0.376$&$0.090$    \\
$C_z$&0.124&$0.084$&$0.105$\\
\hline
\end{tabular*}
%\end{minipage}
\end{table}

>>>>

\begin{abstract}
\ldots 
\end{abstract}

\begin{keywords}
\ldots
\end{keywords}

@@

%FAG rephrase
The decomposition of the magnetic field into mean and fluctuating (random) 
field follows \MHD.
%end FAG
%When considering the mean-field decomposition of the magnetic field, the notation used by \citep{GSSFM13} is adopted.
Volume averaging with a Gaussian kernel, $G_{l}(\bvec{x}-\bvec{x}')$ is used to decompose the magnetic field, $\bvec{B}$, into mean (large-scale), $\bvec{B}_{l}$ and random (small-scale), $\bvec{b}_{l}$, fields. The decomposition is given by 
\begin{equation}
\bvec{B} = \bvec{B}_{l}+\bvec{b}_{l},\qquad \bvec{B}_{l} = \avg{B}. \label{decomp} 
\end{equation}
Gaussian smoothing of the magnetic field with kernel size $l$ is represented by
\begin{align}
\avg{B}&=\int_{V}\bvec{B}(\bvec{x}')G_{l}(\bvec{x}-\bvec{x}')d^{3}\bvec{x}',\\
G_{l}(\bvec{x})&=(2\pi l^2)^{-3/2}\exp[-\bvec{x}/(2l^2)],\nonumber
\label{gauss}
\end{align} 
where $l\approx50$pc is half the integral scale of random motions, as discussed
in \HD.  
Preliminary analysis does not show significant sensitivity of the mean or random field to variations in $l$ within the range $30<l<100$pc. However, future work will consider analysis of isolated structures in the ISM. A more detailed analysis of length and correlation scales will be carried out. However, the current definition, $l=50$pc, is sufficient for the purposes of this paper. 
%\begin{table*}
 %\centering
 %\begin{minipage}{140mm}
  %\caption{Data on the RV Tauri stars detected by {\it IRAS}.}
  %\begin{tabular}{@{}llrrrrlrlr@{}}
  %\hline
  % Name     &            & \multicolumn{4}{c}{Flux density (Jy)%
 %\footnote{Observed by {\em IRAS}.}}\\
 % Variable & {\it IRAS} & 12$\,\umu$m & 25$\,\umu$m & 60$\,\umu$m
 %    & 100$\,\umu$m & Sp. & Period & Light- & $T_0\,(\rmn{K})$ \\
 %       &  &  &  &  &  & group & (d) & curve \\
 %       &  &  &  &  &  &       &     & type  \\
 %\hline
 %TW Cam & 04166$+$5719 & 8.27 & 5.62 & 1.82 & $<$1.73 & A & 85.6 & a & 555 \\
 %RV Tau & 04440$+$2605 & 22.53 & 18.08 & 6.40 & 2.52 & A & 78.9 & b & 460 \\
%\hline
%\end{tabular}
%\end{minipage}
%\end{table*}

@@

  The effect of the magnetic pressure in Model~$\Ompa$ is to 
  expand the thick disc further than Model~$\Omph$ and this is illustrated in 
  Fig.~\ref{fig:zrhom}a, where the horizontal averages of gas number density
  $n(z)$ are plotted against $z$ for both models. 
  The strong peak in the density at the mid-plane is evident for Model~$\Omph$
  (black, solid), while for Model~$\Ompa$ (blue, dashed) there is a broad
  plateau in the density, extending to $|z|\simeq300\pc$, where the mean 
  magnetic field is strongest.

@@

  The horizontal averages of the pressure are plotted in Fig.~\ref{fig:zrhom}b
  for both models. 
  There is a strong peak at the mid-plane in the turbulent pressure for the HD
  model (black, dashed), but a much weaker profile for the MHD model (blue,
  dash-dotted). 
  There are two peaks in the magnetic pressure (blue, dash-3dotted) near 
  $|z|\simeq200\pc$, which supports the extended density profile.
  Another possible effect, which might constrain the circulation of the hot
  gas, and hence enhance the pressure at the mid-plane, is the strong 
  horizontal orientation of the field.
  As mentioned in Section~$7.2$ in \citet{Gent12} the periodic boundary conditions
  exclude a non-zero vertical component to the mean field, so the magnetic
  tension predominantly acts against the vertical flows.
  Some understanding of the multi-phase structure of the magnetised ISM is
  still possible from these models, but the extreme temperatures and densities
  are significantly under represented.
  To improve this in future work it will be desirable to allow unrestricted 
  evolution of vertical field and to apply realistic clustering of the 
  SNe to generate more supperbubbles (composite multiple SN remnants forming a
  single superstructure) or chimneys (plumes venting hot gas from the disc 
  towards the halo).

@@

Even though the visualisation is informative, it is necessary to determine whether the field lines are sensitive to the phases of the ISM. Consequently, we compare the characteristics of the field lines against straight lines launched in the azimuthal ($y$) direction, from the seed points used to compute the field lines. This test is devised using ideas from the field of stereology \citep{BJ04}, lineal analysis in particular. The straight lines used in the test case will have the same general characteristics as the field lines, since they both traverse the computational domain in the positive $y$-direction. This ensures that the lines are representative of the field lines being sampled. Despite having the same general characteristics, the straight lines should not be sensitive to the phases of the ISM; they will not prefer to stay in a particular phase, whilst avoiding others. Therefore, comparison of the PDFs of the observables sampled along the field lines and the straight lines, should identify sensitivity (or lack thereof) of the field lines to the phases of the ISM.\\

@@

The mean magnetic field is considered first, to explore the large-scale characteristics of the magnetic field, having extracted the small-scale fluctuations with Gaussian volume averaging.  Intuition suggests that the mean magnetic field will prefer to stay in the warm phase, which provides a less hostile environment for the magnetic field than the transonic, compressible turbulence of the hot phase. As discussed in Section \label{sect:3BB}, the magnetic field is very strong and well-aligned with the azimuthal ($y$) direction in the $|z|<0.6$kpc region, where the warm phase dominates. In the $0.6<|z|<1.0$kpc region, the hot phase dominates. Consequently, the magnetic field is not well aligned and much weaker than the magnetic field in the warm phase. The the turbulent flows in the $0.6<|z|<1.0$kpc region will have the highest kinetic energies throughout the ISM (CALCULATE THESE AVERAGE VALUES). Conversely, the magnetic field is very weak in the hot phase (Fig.~\ref{fig:b2dnt}). Subsequently, the kinetic energy of the flows should be at least of a comparable order to the magnetic field energy in this region.  We expect this to cause distortions in the magnetic field (CCE: REFERENCE TAYLOR HERE). The lack of alignment and weakness of the magnetic field in this region (Fig. \ref{fig:b3box}) imply that the magnetic field is sensitive to the hot phase. The characteristics of the magnetic field lines further clarify the effect of the hot phase on the magnetic field.

@@

In Figure \ref{fig:fld_lines}, data from the snapshot at $t=1.4$ Gyr are used. The white field lines have been seeded along the $xz$-axis at $y=-0.51$ kpc. The translucent gas inside the box is a representation of the log density field. Density decreases from brighter to darker colours.  
Since the warm and hot phase represents approximately $\%99$ of the fractional
volume of the ISM gas, it is only useful for visualising the characteristics 
of the field line within the warm and hot phases. 
We will present a preliminary result regarding the field lines within the 
cold phase in this paper. 
However, a more detailed discussion will be reported elsewhere. 
For descriptive purposes, the yellow shades of gas correspond to warm gas and 
the red shades correspond to the hot phase. 
We note that the field lines are generally aligned with the positive 
$y$-direction, especially in the warm phase and the field lines are smooth 
inside the warm gas. 
However, we notice significant differences in the characteristics of the 
field lines in the hot phase.

@@

Similarly, in the range $0.7<z<1.1$ kpc, there is a larger structure of hot 
gas, which spans the $xy$-plane. 
This structure appears to be part of rising hot gas, which is leaving the 
computational domain. As it leaves, it elongates the magnetic field lines 
until they are no longer connected inside the domain. 
Consequently, the the field lines are approximately perpendicular to the 
$y$-axis, where the magnetic field lines are strecthed and distorted by the 
hot phase.
\noindent Figure \ref{fig:pdfs_mean} shows the probability density functions (PDFs) of 
entropy sampled along field lines (solid lines) and straight lines (dashed
lines) for the snapshot at $t=1.4$ Gyr. 
The PDFs show that the magnetic field lines and straight lines have similar 
overall behaviour, due to their alignment in the $y$-direction. 
The alignment of the straight lines in the azimuthal ($y$) direction was 
chosen to ensure that the straight lines are representative of the general 
alignment of the magnetic field, without introducing any sensitivity to the 
multi-phase structure.

@@

Overall, the magnetic field lines traverse the ISM in the azimuthal direction. 
This trend is observed clearly in the warm phase, where the magnetic field is 
not distorted by the flows of the ISM gas. 
Since the field lines represent lines of magnetic flux density, the 
distortion of the magnetic field lines by the hot phase explains the weakness 
of the magnetic field observed in the hot phase. 
The stretching of the magnetic field lines effectively wrap the field lines 
around the hot gas, which increases the field line density on the boundary 
of the hot and warm gas. 
Consequently, field lines do not pass through the hot gas 
(Fig.~\ref{fig:fld_lines}) and we observe that the magnetic field in the hot 
phase is very weak and lacking alignment. 
This strongly suggests that the magnetic field resides in the warm phase.    
%-----------------------------------------------------------------------------
\begin{table}
 \centering
 %\begin{minipage}{180mm}
  \caption{Respective probabilities of finding the field and straight in the cold, warm and hot phases.}
  \label{table:mean_probs}
 % \begin{tabular*}{@{}c c c @{}}{\linewidth}{@{\extracolsep{\fill}}p{0.3\linewidth}p{0.3\linewidth}p{0.3\linewidth}@{}}
\begin{tabular*}{\linewidth}{@{\extracolsep{\fill}}p{0.3\linewidth}p{0.3\linewidth}p{0.3\linewidth}@{}}
  \hline
   Probability& Mean field lines& Straight lines \\
 \hline
\\
$\mathcal P(\text{cold phase})$&$\ll0.01$ &$\ll0.01$  \\
   $\mathcal P(\text{warm phase})$& $0.79$&$0.89$   \\
   $\mathcal P(\text{hot phase})$&$0.20$ & $0.11$    \\
\hline
\end{tabular*}
%\end{minipage}
\end{table}
%-----------------------------------------------------------------------------

@@

%-----------------------------------------------------------------------------
%\subsection{Curve fitting for PDFs of entropy}
%-----------------------------------------------------------------------------
\subsection{Correlation lengths of the magnetic fields}
\begin{table}[H]
 \centering
 %\begin{minipage}{180mm}
  \caption{Correlation lengths for the total, mean and random magnetic field in kpc.}
  \label{table:mean_probs}
 % \begin{tabular*}{@{}c c c @{}}{\linewidth}{@{\extracolsep{\fill}}p{0.3\linewidth}p{0.3\linewidth}p{0.3\linewidth}@{}}
\begin{tabular*}{\linewidth}{p{0.20\linewidth}p{0.20\linewidth}p{0.2\linewidth}p{0.2\linewidth}@{}}
  \hline
Correlation length &Total magnetic field& Mean magnetic field& Random magnetic field \\
 \hline
\\
$C$&$0.711$&$0.700$&$0.241$ \\
$C_x$& $0.061$&$0.045$&$0.065$   \\
$C_y$&$0.380$ & $0.376$&$0.090$    \\
$C_z$&0.124&$0.084$&$0.105$\\
\hline
\end{tabular*}
%\end{minipage}
\end{table}
